---
title: "Chapter 5. Dimensionality reduction techniques"
author: "Tuure Parviainen"
date: "December 1, 2017"
output: html_document
---
## Chapter 5. Dimensionality reduction techniques



Today we are looking at 

### Boston Data set information:

Variables     | Description
----------    | --------------------------------------
crim          | per capita crime rate by town.
zn            | proportion of residential land zoned for lots over 25,000 sq.ft.
indus         | proportion of non-retail business acres per town.
chas          | Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).
nox           | nitrogen oxides concentration (parts per 10 million).
rm            | average number of rooms per dwelling.
age           | proportion of owner-occupied units built prior to 1940.
dis           | weighted mean of distances to five Boston employment centres.
rad           | index of accessibility to radial highways.
tax           | full-value property-tax rate per \$10,000.
ptratio       | pupil-teacher ratio by town.
black         | 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.
lstat         | lower status of the population (percent).
medv          | median value of owner-occupied homes divided by $1000s.


```{r, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
require("easypackages") # for loading and installing many packages at one time
packages_to_load <- c("broom", "dplyr", "MASS", "tidyverse", "corrplot", "ggplot2", "GGally", "caret", "devtools", "ggthemes", "scales", "plotly")
packages(packages_to_load, prompt = TRUE) # lock'n'load install/load
#Additionally
#install_github("fawda123/ggord") # Installed from Github for vizualization
library(ggord)

```

```{r, message=FALSE, warning=FALSE}

# To empty the memory after the excercise before this
# rm(list=ls())
# Load data
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # to set working directory to source file

# load the data
data("Boston")
glimpse(Boston)

```


