---
title: "Chapter 4. Clustering and classification"
author: "Tuure Parviainen"
date: "November 22, 2017"
output: html_document
---
## Chapter 4. Clustering and classification

I completed all the DataCamp exercises prior to proceeding to this exercise. I also looked up some details how to use the RMarkdown more productively and hopefully this report will be clearer than the previous ones.

Today we are looking at the Boston crime

### Boston Data set information:

Variables     | Description
----------    | --------------------------------------
crim          | per capita crime rate by town.
zn            | proportion of residential land zoned for lots over 25,000 sq.ft.
indus         | proportion of non-retail business acres per town.
chas          | Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).
nox           | nitrogen oxides concentration (parts per 10 million).
rm            | average number of rooms per dwelling.
age           | proportion of owner-occupied units built prior to 1940.
dis           | weighted mean of distances to five Boston employment centres.
rad           | index of accessibility to radial highways.
tax           | full-value property-tax rate per \$10,000.
ptratio       | pupil-teacher ratio by town.
black         | 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.
lstat         | lower status of the population (percent).
medv          | median value of owner-occupied homes divided by $1000s.


```{r, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
require("easypackages") # for loading and installing many packages at one time
packages_to_load <- c("broom", "dplyr", "MASS", "tidyverse", "corrplot", "ggplot2", "GGally", "caret", "devtools", "ggthemes", "scales", "plotly")
packages(packages_to_load, prompt = TRUE) # lock'n'load install/load
#Additionally
#install_github("fawda123/ggord") # Installed from Github for vizualization
library(ggord)

```

```{r, message=FALSE, warning=FALSE}

# To empty the memory after the excercise before this
# rm(list=ls())
# Load data
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # to set working directory to source file

# load the data
data("Boston")
glimpse(Boston)

```

All are numeric variables except chas and rad. The data has demographical data of boston including tax and other information possibly linked to crime rates within the city.

```{r, message=FALSE, warning=FALSE}

my_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) +
    geom_point() +
    geom_smooth(method=loess, fill="red", color="red", ...) +
    geom_smooth(method=lm, fill="blue", color="blue", ...)
  p
}
g = ggpairs(Boston,columns = c(1:14), lower = list(continuous = my_fn))
g

```

1. It seems that indus, nox, rad, tax and lstat have positive correlation with crime, and dis, black and medv have negative correlation with crime.
2. Many of the variables seems to have non-linear relationships (red lines in the picture), however the variables with high correlation seems to be more linear.

```{r, message=FALSE, warning=FALSE}

# calculate the correlation matrix and round it
cor_matrix<-cor(Boston) %>% round(digits = 2)

# visualize the correlation matrix
corrplot(cor_matrix, method="circle", type = "upper", cl.pos = "b", tl.pos = "d", tl.cex = 0.6)

```

More focused figure on correlation, since they are hardly visible with this many variables.

```{r, message=FALSE, warning=FALSE}

# Scaling the variables with mean and standard deviation with scale()
Boston_scaled <- as.data.frame(scale(Boston))
# create a quantile vector of crime and print it
bins <- quantile(Boston_scaled$crim)
bins
# create a categorical variable 'crim'
crime <- cut(Boston_scaled$crim, breaks = bins, include.lowest= TRUE, label = c("low","med_low","med_high", "high"))

# remove original crim from the dataset
Boston_scaled <- dplyr::select(Boston_scaled, -crim)

# add the new categorical value to scaled data
Boston_scaled <- data.frame(Boston_scaled, crime)

# number of rows in the Boston dataset
n <-nrow(Boston)

# choose randomly 80% of the rows
ind <- sample(n,  size = n * 0.8)

# create train set
train <- Boston_scaled[ind,]

# create test set
test <- Boston_scaled[-ind,]

# save the correct classes from test data
correct_classes <-test$crime

# remove the crime variable from test data
test <- dplyr::select(test, -crime)

summary(train)

```

Now we have generated the test dataset with the new categorial variable "crime", which is based on the quantiles of the original numeric variable.

```{r, message=FALSE, warning=FALSE}

# linear discriminant analysis
lda.fit <- lda(crime ~., data = train)

# print the lda.fit object
lda.fit

g <- ggord(lda.fit, train$crime, ellipse_pro = 0.95, vec_ext = 2, alpha_el = 0.3, size = 2)
g  + theme_tufte()  + geom_rangeframe()

```

1. We get the points seperated only by the 1st LD and rad parameter in the high group.
2. From the LDA summary, we can see that how the first linear discriminant explain 94.72 % of the between-group variance in the Boston dataset.
3. Also the rad (= index of accessibility to radial highways) has relatively high value in the first discriminant, so it proves to be uselful in the classification. rad had negative correlation with crime.

```{r, message=FALSE, warning=FALSE}
# predict classes with test data
lda.pred <- predict(lda.fit, newdata = test)

# cross tabulate the results
table(correct = correct_classes, predicted = lda.pred$class)

```

In the high -crime quartile, we could predict all cases, but for other classes we didn't do so well. However, LDA seems to be able to predict our cases with quite good accuracy.

```{r, message=FALSE, warning=FALSE}
# k-means clustering
BostonS <- scale(Boston) # standardize variables

wss <- (nrow(BostonS)-1)*sum(apply(BostonS,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(BostonS, 
  	centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")


km <-kmeans(BostonS, centers = 4)

# plot the Boston dataset with clusters
pairs(BostonS, col = km$cluster)
```
We see similar results than in LDA

```{r, message=FALSE, warning=FALSE}
model_predictors <- dplyr::select(train, -crime)
# check the dimensions
dim(model_predictors)
dim(lda.fit$scaling)
# matrix multiplication
matrix_product <- as.matrix(model_predictors) %*% lda.fit$scaling
matrix_product <- as.data.frame(matrix_product)

plot_ly(x = matrix_product$LD1, y = matrix_product$LD2, z = matrix_product$LD3, type= 'scatter3d', mode='markers')

```

Now we can see the 3D picture of the clusters!